\paperTitle

### 2014-05-14 doc/icml2014/directed.tex
fix lemma 2

% DONE: don't talk about parameter recovery yet, just establish exclusive views
%Finally, the claim we make is that we can recover parameters for any
  %graphical model where every hidden variable is a bottleneck:
Now define the following property:
\begin{property}[Uniformly bottlenecked]
  \label{prop:bottleneck}
  A clique $\sC$ is \emph{uniformly bottlenecked} if
  every hidden variable in $\sC$ is a bottleneck. 
  If every clique $\sC \in \sG$ is uniformly bottlenecked, then we say
  that the graph $\sG$ is uniformly bottlenecked as well.
\end{property}
Note that two hidden variables are allowed to have overlapping views.
%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
It is clear that having uniform bottlenecks implies part (ii)
of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
but we will show that part (i) follows as well, given that the graphical model faithfully represents conditional independences:

\begin{assumption}[Faithful Representation]
  \label{asm:ci}  
  If there is an active trail from any two variables $a, b$ in the graph
    $\sG$, then they are not independent, i.e $a \not\perp b$. 
  If there is an active trail between $a$ and $b$ conditioned on any variable,
    say $c$, then they are also not independent, i.e. $a\not\perp b \given c$.
\end{assumption}

@@

%Before we establish this relationship,
First, we describe a reduction of any graphical model to
  a canonical form to make our arguments simpler.
  % DONE: don't say this yet since reader can't appreciate it.
  %that will allow us to handle the case where an observed
  %variable has multiple parents.

@@

Now we establish the link between bottlenecks and exclusive views
(see \appendixref{exclusive-views} for the complete proof):
\begin{lemma}[Uniformly bottlenecked implies exclusive views]
  \label{lem:bottleneck-views}  
% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
  %a connected subgraph of hidden variables.
  %If every hidden variable
  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
  If a graphical model $\sG$ satisfies the faithful representation 
  assumption (\assumptionref{ci}) and is uniformly bottlenecked, then
  $\sG$ has the exclusive views property.
\end{lemma}
% PL: save space, doesn't add much
%\begin{proof}
%  The independence conditions required by the bottleneck condition
%  effectively guarantee that at least one of the three views for
%  a bottleneck be exclusive. 
%  \appendixref{exclusive-views} provides a complete proof.
%\end{proof}

>>>>

% DONE: don't talk about parameter recovery yet, just establish exclusive views
%Finally, the claim we make is that we can recover parameters for any
  %graphical model where every hidden variable is a bottleneck:
Now define the following property:
\begin{property}[Uniformly bottlenecked]
  \label{prop:bottleneck}
  A clique $\sC$ is \emph{uniformly bottlenecked} if
  every hidden variable in $\sC$ is a bottleneck. 
  If every clique $\sC \in \sG$ is uniformly bottlenecked, then we say
  that the graph $\sG$ is uniformly bottlenecked as well.
\end{property}
Note that two hidden variables are allowed to have overlapping views.
%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
It is clear that having uniform bottlenecks implies part (ii)
of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
but we will show that part (i) follows as well.

To rule out edge cases, we assume the graphical model only has conditional
independences given by the graph:
\begin{property}[Faithful representation]
  \label{prop:ci}  
  %If there is an active trail from any two variables $a, b$ in the graph
    %$\sG$, then they are not independent, i.e $a \not\perp b$. 
  We say that a graphical model $\sG$ is \emph{faithful}
  if for any variables $a,b,c$
  such that there is an active trail\footnote{See \todo{cite textbook} for a definition.} between $a$ and $b$ conditioned on $c$,
  then $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
\end{property}

@@

%Before we establish this relationship,
We now describe a reduction of any graphical model to
  a canonical form to make our arguments simpler.
  % DONE: don't say this yet since reader can't appreciate it.
  %that will allow us to handle the case where an observed
  %variable has multiple parents.

@@

Now we establish the link between bottlenecks and exclusive views
(see \appendixref{exclusive-views} for the complete proof):
\begin{lemma}[Uniformly bottlenecked implies exclusive views]
  \label{lem:bottleneck-views}  
% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
  %a connected subgraph of hidden variables.
  %If every hidden variable
  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
  If a graphical model $\sG$ is faithful and uniformly bottlenecked, then
  $\sG$ has the exclusive views property.
\end{lemma}
\begin{proof}
We prove the contrapositive: suppose that there is some hidden clique $\sC \in \sG$ with $h_0 \in \sC$
that does not have exclusive views.  This means each observed variable $x_i$ is conditionally dependent on some
$h_i \in \sC \backslash \{h_0\}$ given $h_0$.
Consider any such two $x_1, x_2$.
Conditioned on $h_0$, there exists an active trail between them (specifically, $x_1 - h_1 - h_2 - x_2$).
Since $\sG$ is faithful, we have $x_1 \not\perp x_2 \mid h_0$.
As this holds for any $x_1,x_2$, $h_0$ cannot be a bottleneck.
\end{proof}
